Chord: A Versatile Platform for Program Analysis

### 2011-03-15 main/doc/dynamic_analysis.tex
more documentation

A sample such class \code{MyDynamicAnalysis} is shown below:

\begin{verbatim}
    import chord.project.analyses.DynamicAnalysis;
    import chord.instr.InstrScheme;

    // ***TODO***: analysis won't be recognized by Chord without this annotation
    @Chord(name=<name-of-analysis>)    
    public class MyDynamicAnalysis extends DynamicAnalysis {
        InstrScheme scheme;

        @Override
        public InstrScheme getInstrScheme() {
            if (scheme != null)
                return scheme;
            scheme = new InstrScheme();
            // ***TODO***: Choose (<event1>, <args1>), ... (<eventN>, <argsN>)
            // depending upon the kind and format of events required by this
            // dynamic analysis to be generated for this during an instrumented
            // program's execution.
            scheme.set<event1>(<args1>);
            ...
            scheme.set<eventN>(<argsN>);
            return scheme;
        }

        @Override
        public void initAllPasses() {
            // ***TODO***: User code to be executed once and for all
            // before all instrumented program runs start.
        }

        @Override
        public void doneAllPasses() {
            // ***TODO***: User code to be executed once and for all
            // after all instrumented program runs finish.
        }

        @Override
        public void initPass() {
            // ***TODO***: User code to be executed once before each
            // instrumented program run starts.
        }

        @Override
        public void donePass() {
            // ***TODO***: User code to be executed once after each
            // instrumented program run finishes.
        }

        // User-defined event handlers for this dynamic analysis.
        // No-ops if not overridden.
        @Override
        public void process<event1>(<args1>) {
            // ***TODO***: User code for handling events of kind
            // <event1> with format <args1>.
        }
        ...
        @Override
        public void process<eventN>(<argsN>) {
            // ***TODO***: User code for handling events of kind
            // <eventN> with format <argsN>.
        }
    }
\end{verbatim}

@@

Before thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. 
{\bf Note:} This event is not generated before constructor calls; use the BefNew event.

@@

After thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. 
{\bf Note:} This event is not generated after constructor calls; use the AftNew event.

>>>>

A sample such class \code{MyDynamicAnalysis} is shown in Figure \ref{fig:dynamic-analysis}.

\begin{figure}
{\small
\begin{verbatim}
    import chord.project.analyses.DynamicAnalysis;
    import chord.instr.InstrScheme;

    // ***TODO***: analysis won't be recognized by Chord without this annotation
    @Chord(name=<name-of-analysis>)    
    public class MyDynamicAnalysis extends DynamicAnalysis {
        InstrScheme scheme;
        @Override
        public InstrScheme getInstrScheme() {
            if (scheme != null)
                return scheme;
            scheme = new InstrScheme();
            // ***TODO***: Choose (<event1>, <args1>), ... (<eventN>, <argsN>)
            // depending upon the kind and format of events required by this
            // dynamic analysis to be generated for this during an instrumented
            // program's execution.
            scheme.set<event1>(<args1>);
            ...
            scheme.set<eventN>(<argsN>);
            return scheme;
        }
        @Override
        public void initAllPasses() {
            // ***TODO***: User code to be executed once and for all
            // before all instrumented program runs start.
        }
        @Override
        public void doneAllPasses() {
            // ***TODO***: User code to be executed once and for all
            // after all instrumented program runs finish.
        }
        @Override
        public void initPass() {
            // ***TODO***: User code to be executed once before each instrumented program run starts.
        }
        @Override
        public void donePass() {
            // ***TODO***: User code to be executed once after each instrumented program run finishes.
        }
        @Override
        public void process<event1>(<args1>) {
            // ***TODO***: User code for handling events of kind <event1> with format <args1>.
        }
        ...
        @Override
        public void process<eventN>(<argsN>) {
            // ***TODO***: User code for handling events of kind <eventN> with format <argsN>.
        }
    }
\end{verbatim}
}
\label{fig:dynamic-analysis}
\caption{Dynamic Analysis Template.}
\end{figure}

@@

\begin{tabular}{p{4.3in}}
Before thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. \\
{\bf Note:} Not generated before constructor calls; use BefNew event.

@@

\begin{tabular}{p{4.3in}}
After thread \bt\ executes the method invocation statement at program point \bi\ (in domain I) with this argument as object \bo. \\
{\bf Note:} Not generated after constructor calls; use AftNew event.

### 2011-03-15 main/doc/example.tex
more documentation

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<PROGRAM DIR NAME> -Dchord.run.analyses=<name1>,...,<nameN> run
\end{verbatim}

where \code{<PROGRAM DIR NAME>} specifies an example Java program as described above,
and each of \code{<name1>}, ..., \code{<nameN>} is the name of an analysis to run,
such as \code{<JAVA ANALYSIS NAME>} and \code{<DLOG ANALYSIS NAME>} in the Java and
Datalog analyses depicted above.

For instance, to run a basic may-alias and call-graph analysis (called 0CFA),
run the following command in Chord's main directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<PROGRAM DIR NAME> -Dchord.run.analyses=cipa-0cfa-dlog run
\end{verbatim}

@@

    run:
         [java] Chord run initiated at: Mar 13, 2011 10:31:08 PM
         [java] ENTER: cipa-0cfa-dlog
         [java] ENTER: T
         [java] ENTER: RTA
         [java] Iteration: 0
         [java] Iteration: 1
         [java] Iteration: 2
         [java] LEAVE: RTA
         [java] SAVING dom T size: 1386
         [java] LEAVE: T
         [java] ENTER: F
         [java] SAVING dom F size: 4120
         [java] LEAVE: F
         ...
         [java] ENTER: MputStatFldInst
         [java] SAVING rel MputStatFldInst size: 739
         [java] LEAVE: MputStatFldInst
         [java] ENTER: statIM
         [java] SAVING rel statIM size: 3359
         [java] LEAVE: statIM
         [java] Starting command: 'java ... net.sf.bddbddb.Solver ...\chord_analyses_alias_cipa_0cfa.dlog'
         [java] Relation VH: 541 nodes, 449.0 elements (V0,H0)
         [java] Relation FH: 137 nodes, 8.0 elements (H0,F0)
         [java] Relation HFH: 199 nodes, 35.0 elements (H0,F0,H1)
         [java] Relation IM: 590 nodes, 69.0 elements (I0,M0)
         [java] Finished command: 'java ... net.sf.bddbddb.Solver ...\chord_analyses_alias_cipa_0cfa.dlog'
         [java] LEAVE: cipa-0cfa-dlog
         [java] Chord run completed at: Mar 13, 2011 10:31:36 PM
         [java] Total time: 00:00:27:671 hh:mm:ss:ms

@@

The program relations consumed by the 0CFA analysis contain basic program facts.  For instance,
\code{MputStatFldInst} is a relation containing each tuple (m,f,v) such that method m in the input Java
program contains a putstatic instruction of the form ``f = v", while \code{statIM} is a relation
containing each tuple (i,m) such that m is the target method of invokestatic instruction i.

The program relations produced by the 0CFA analysis represent points-to information and the
call graph of the input Java program as computed by the analysis.
Specifically, relations \code{VH}, \code{FH}, and \code{HFH} represent points-to information 
for local variables, static fields, and instance fields, respectively,
while relation \code{IM} represents the call graph, namely, it contain each tuple (i,m) such that
m is a possible target method of call site i.

Metavariables m, f, i, and v above range over entities in so-called {\it program domains} M, F, I, and V, respectively.
A program domain is a set of entities of a certain kind in the input Java program.  For instance, M is the domain representing
the set of all methods in the input Java program, F is the domain representing the set of all fields, I is the domain representing
the set of all call sites in methods in M, and V is the domain representing the set of all local variables of reference type in methods in M.
Each of these domains is produced by a separate Java analysis with the corresponding name.
Notice that the analyses producing these domains run upfront because these domains
are consumed by the analyses that produce relations such as \code{MputStatFldInst} and
\code{statIM}, which in turn are consumed by the desired 0CFA analysis.

During execution, Chord dumps intermediate and final results to files in the
directory specified by property \code{chord.out.dir}, whose default value is \code{[chord.work.dir]/chord_output/}
and typically does not need to be changed by users.
For the above example, this directory is \code{<PROGRAM DIR NAME>/chord_output/}.

>>>>

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<PROGRAM DIR> -Dchord.run.analyses=<ANALYSIS NAME> run
\end{verbatim}

where {\tt <PROGRAM DIR>} denotes a directory containing a Java program as described above,
and {\tt <ANALYSIS NAME>} denotes the name of an analysis to run,
such as {\tt <JAVA ANALYSIS NAME>} and {\tt <DLOG ANALYSIS NAME>} in the Java and
Datalog analyses depicted above.

For instance, to run a basic may-alias and call-graph analysis (called 0CFA),
run the following command in Chord's \code{main/} directory:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<PROGRAM DIR> -Dchord.run.analyses=cipa-0cfa-dlog run
\end{verbatim}

@@

    run:
         [java] Chord run initiated at: Mar 13, 2011 10:31:08 PM
         [java] ENTER: cipa-0cfa-dlog
         [java] ENTER: T
         [java] ENTER: RTA
         [java] Iteration: 0
         [java] Iteration: 1
         [java] Iteration: 2
         [java] LEAVE: RTA
         [java] SAVING dom T size: 1386
         [java] LEAVE: T
         [java] ENTER: F
         [java] SAVING dom F size: 4120
         [java] LEAVE: F
         ...
         [java] ENTER: MputStatFldInst
         [java] SAVING rel MputStatFldInst size: 739
         [java] LEAVE: MputStatFldInst
         [java] ENTER: statIM
         [java] SAVING rel statIM size: 3359
         [java] LEAVE: statIM
         [java] Starting command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
         [java] Relation VH: 541 nodes, 449.0 elements (V0,H0)
         [java] Relation FH: 137 nodes, 8.0 elements (H0,F0)
         [java] Relation HFH: 199 nodes, 35.0 elements (H0,F0,H1)
         [java] Relation IM: 590 nodes, 69.0 elements (I0,M0)
         [java] Finished command: 'java ... chord_analyses_alias_cipa_0cfa.dlog'
         [java] LEAVE: cipa-0cfa-dlog
         [java] Chord run completed at: Mar 13, 2011 10:31:36 PM
         [java] Total time: 00:00:27:671 hh:mm:ss:ms

@@

The program relations consumed by the 0CFA analysis contain basic program facts.  For instance,
\code{MputStatFldInst} is a relation containing each tuple ($m$,$f$,$v$) such that method $m$ in the input Java
program contains a putstatic instruction of the form ``$f$ = $v$", while \code{statIM} is a relation
containing each tuple ($i$,$m$) such that $m$ is the target method of invokestatic instruction $i$.

The program relations produced by the 0CFA analysis represent points-to information and the
call graph of the input Java program as computed by the analysis.
Specifically, relations \code{VH}, \code{FH}, and \code{HFH} represent points-to information 
for local variables, static fields, and instance fields, respectively,
while relation \code{IM} represents the call graph, namely, it contain each tuple ($i$,$m$) such that
$m$ is a possible target method of call site $i$.

Metavariables $m$, $f$, $i$, and $v$ above range over entities in so-called {\it program domains} \code{M},
\code{F}, \code{I}, and \code{V}, respectively.
A program domain is a set of entities of a certain kind in the input Java program.  For instance,
\code{M} is the domain representing
the set of all methods in the input Java program, \code{F} is the domain representing the set of all fields,
\code{I} is the domain representing
the set of all call sites in methods in \code{M}, and \code{V} is the domain representing the set of all local variables of reference type in methods in \code{M}.
Each of these domains is produced by a separate Java analysis with the corresponding name.
Notice that the analyses producing these domains run upfront because these domains
are consumed by the analyses that produce relations such as \code{MputStatFldInst} and
\code{statIM}, which in turn are consumed by the desired 0CFA analysis.

During execution, Chord dumps intermediate and final results to files in the
directory specified by property \code{chord.out.dir}, whose default value is \code{[chord.work.dir]/chord_output/}
and typically does not need to be changed by users.
For the above example, this directory is {\tt <PROGRAM DIR>}\code{/chord_output/}.

### 2011-03-15 main/doc/properties.tex
more documentation

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...> run
\end{verbatim}
where ``\code{<...>}" denotes the absolute or relative path of the top-level directory of
the program to be analyzed, which contains a file named \code{chord.properties} that users
must create themselves and populate with properties that are recognized by chord and
are specific to that program.

@@

This section describes properties of the Java program to be analyzed, such as
its main class, the location(s) of its class files and Java source
files, and command-line arguments to be used to run the program.
\\[10pt]

@@

This section describes properties that specify how the analysis scope of the input Java program is computed.
See Chapter \ref{chap:computing-scope} for more details.
\\[10pt]

@@

This section describes properties that dictate what task(s) Chord must perform.
\\[10pt]

@@

This section describes properties regarding analyses executed by Chord.
\\[10pt]

@@

This section describes properties regarding bytecode instrumentation and dynamic analysis.
\\[10pt]

@@

This section describes properties that specify what must be reused by Chord, if available, from previous runs instead of recomputing.
\\[10pt]

@@

This section describes properties regarding the JVM that runs Chord.
\\[10pt]

@@

This section describes properties concerning BDD-based Datalog solver bddbddb that is used by Chord to run analyses written in Datalog.
\\[10pt]

@@

This section describes properties specifying the names of files and directories output by Chord.
Most users will not need to alter the default values of these properties.
\\[10pt]

>>>>

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...> run
\end{verbatim}
where \code{<...>} denotes the absolute or relative path of the top-level directory of
the program to be analyzed, which contains a file named \code{chord.properties} that users
must create themselves and populate with properties that are recognized by chord and
are specific to that program.

@@

This section describes properties of the Java program to be analyzed, such as
its main class, the location(s) of its class files and Java source
files, and command-line arguments to be used when running the program.

@@

This section describes properties that specify how the analysis scope of the input Java program is computed.
See Chapter \ref{chap:computing-scope} for more details.

@@

This section describes properties that dictate what task(s) Chord must perform.

@@

This section describes properties regarding analyses executed by Chord.

@@

This section describes properties regarding bytecode instrumentation and dynamic analysis.

@@

This section describes properties that specify what must be reused by Chord, if available, from previous runs instead of recomputing.

@@

This section describes properties regarding the JVM that runs Chord.

@@

This section describes properties concerning BDD-based Datalog solver bddbddb that is used by Chord to run analyses written in Datalog.

@@

This section describes properties specifying the names of files and directories output by Chord.
Most users will not need to alter the default values of these properties.
